Don't Mind the Gap: Bridging Network-wide Objectives and Device-level Configurations

### 2016-01-28 paper/introduction.tex
tweaks to abstract and intro

It is well-known that network configuration is highly error
prone and that such errors can lead to costly network
downtime~\cite{mahajan+:bgp-misconfiguration,feamster+:rcc,batfish}.
For instance, a recent misconfiguration led to an hour-long, nation-wide outage for Time Warner's backbone network~\cite{time-warner}; and a major BGP-related incident makes international news every few months~\cite{bgpmon}.

@@

Besides the languages being low-level, a fundamental reason for configuration being onerous is the semantic mismatch between the intended high-level
policies and the low-level device-by-device configurations.  In particular,
many policies involve network-wide properties---prefer a certain neighbor,
never announce a particular destination externally,
use a particular path only if another fails---but configurations describe the behavior of
individual devices.
%
Operators must manually decompose network-wide policy into device-level behaviors, such that their interactions satisfy the high-level objectives.
%
Ensuring policy-compliance through this manual decomposition is challenging under normal
circumstances and even more so in the face of failures. Some decompositions that work correctly
in failure-free environments, violate key constraints when failures occur.
%
As a result, many network configuration problems are only revealed after failures~\cite{batfish},
sometimes making a bad situation worse.

To reduce configuration errors, many practitioners have a adopted a template-based
approach~\cite{hatch,thwack}, in which common tasks are captured as parameterized templates.
%
%More powerful still are systems like
%ConfigAssure~\cite{narain:lisa05,narain+:configassure}, which use SAT solving
%and model finding tools to fill in parameters in configurations while
%ensuring key correctness constraints are satisfied.
One step further, systems like
ConfigAssure~\cite{narain:lisa05,narain+:configassure} use SAT solving
and model finding tools to correctly and consistently fill in some parameters.
%
However,
 while templates help ensure certain kinds of consistency across devices,
they do not bridge the
semantic divide between network policies and device-level configuration.
They do not provide fundamentally different abstractions
from existing configuration languages or new, higher-level reasoning
principles.

@@

The centralized control planes of SDN, however, are not a panacea.
First, while many SDN systems provide effective \emph{intra}-domain routing
abstractions, letting users specify paths within their network,
they fail to provide a coherent means to specify preferred
\emph{inter}-domain routes.  Second, centralized control planes
require careful design and engineering to be robust to failures---one must ensure that all devices can communicate with the controller at all times, even under arbitrary combinations of failure. Even ignoring failures, it is necessary for controller systems to
scale to meet the demands of large, geographically distributed networks,
and to react quickly
to a changing environment. To address this challenge, researchers are exploring
multi-controller systems with interacting controllers, thus bringing back distributed
control planes~\cite{mccauley2013extending,onos} and their current programming difficulties.
%However, academic
%language design and implementation efforts have not kept pace.  For instance, work on many
%experimental SDN languages~\cite{frenetic,flowlog,vericon,merlin,netkat,kinetic,pga} has not yet shown how to implement fault tolerant
%multi-controller systems that support their high-level abstractions efficiently.

Hence, in this paper, we have two central goals:
\begin{enumerate}
\item Design a new, high-leve language with abstractions
for expressing intra-domain routing, inter-domain
routing and routing alternatives in case of failures.
\item Define algorithms for compiling these specifications in to
configurations for devices running standard
distributed control plane algorithms, while ensuring correct behavior
independent of the number of faults.
\end{enumerate}

To achieve the first goal, we borrowed heavily from 
recent high-level SDN languages such as FatTire~\cite{fattire},
Merlin~\cite{foster:merlin}, and
NetKAT~\cite{netkat} by using regular expressions to specify network paths.  However, our design also contains several key
departures from existing languages.  The most important difference is semantic:  the paths specified
by programmers can extend from outside the operator's network to inside
the network, across several devices internally, and then out again.  This design
choice allows users to specify preferences about both external and internal
routes in the same notation.  
%No SDN programming language provides similar facilities.  
In addition,
we incorporate user preferences in to the algebra
of regular expressions and provide a semantics in terms of sets of ranked paths.  The preferences indicate fail-over behaviors:  among all specified paths that are still available,
the system guarantees that the highest-ranked one will be taken.  
%We also support network operators by defining a
%set of common abbreviations for entering, leaving, and traversing
%the user network and we provide abstractions for managing common features %of
%control plane algorithms such as route aggregation.  Operators use familiar logical operations
%to build routes in a modular and compositional manner from these
%primitives.  Our abbreviations are translated in to our regular-expression-based
%intermediate representation and compiled from there.

To achieve the second goal, we develop program analysis and compilation
algorithms that translate the regular policies to a graph-based
intermediate language and from there to BGP configurations.  
We target BGP for pragmatic reasons:  it is an industry
standard and many networks use it internally as well as externally. Despite the advent of SDN, many networks will continue to
use BGP for the foreseeable future due to existing infrastructure investments, the difficulty of transitioning to SDN,
and the scalability and fault-tolerance advantages of a distributed
control plane.

However, using BGP as a vehicle for implementing high-level policies is
challenging because not all policies can be implemented correctly in
the presence of arbitrary failures.  We develop new
algorithms to detect policies that cannot be implemented
and report our findings to the operators, so they may fix the specification at compile time rather than experience undesirable behavior after configurations are deployed.
%We also analyze route aggregation to detect the possibility of failure-induced black-holes.
If the policies are implementable, our compiler automatically synthesizes per-device import and export filters, local preferences, MED attributes, and community tags.

>>>>

It is well known that network configuration is highly error
prone and that such errors can lead to costly network
downtime~\cite{mahajan+:bgp-misconfiguration,feamster+:rcc,batfish}.
For instance, a recent misconfiguration led to an hour-long, nation-wide outage for Time Warner's backbone network~\cite{time-warner}; and a major BGP-related incident makes international news every few months~\cite{bgpmon}.

@@

A fundamental reason for the prevalence of misconfigurations is the
semantic mismatch between the intended high-level
policies and the low-level configurations.  In particular,
many policies involve network-wide properties---prefer a certain neighbor,
never announce a particular destination externally,
use a particular path only if another fails---but configurations describe the behavior of
individual devices.
%
Operators therefore must manually decompose network-wide policy into
device-level behaviors, such that the distributed interactions among
these devices satisfy the high-level objectives.
%
Further, policy-compliance must be ensured not only under normal network
circumstances but also in the face of failures.  The need to reason
about possible failures exacerbates an already considerable challenge
for network operators.  As a result, configurations that work
correctly in failure-free environments have nonetheless been found to violate key
network-wide properties when failures occur~\cite{batfish}.

To reduce configuration errors, many practitioners have a adopted a template-based
approach~\cite{hatch,thwack}, in which common tasks are captured as parameterized templates.
%
%More powerful still are systems like
%ConfigAssure~\cite{narain:lisa05,narain+:configassure}, which use SAT solving
%and model finding tools to fill in parameters in configurations while
%ensuring key correctness constraints are satisfied.
One step further, systems like
ConfigAssure~\cite{narain:lisa05,narain+:configassure} use SAT solving
and model finding tools to correctly and consistently fill in some parameters.
%
While templates help ensure certain kinds of consistency across devices,
they do not bridge the
semantic divide between network-wide policies and device-level configuration.
They do not provide fundamentally different abstractions
from existing configuration languages and so still require operators to
manually decompose policies into per-device behaviors.

@@

The centralized control planes of SDN, however, are not a panacea.
First, while many SDN systems provide effective \emph{intra}-domain routing
abstractions, letting users specify paths within their network,
they fail to provide a coherent means to specify preferred
\emph{inter}-domain routes.  \todd{What is the impact of these
  languages not supporting inter-domain routing?  How are those
  policies expressed instead in an SDN context?  A sentence on the
  impact would help.}  Second, centralized control planes
require careful design and engineering to be robust to failures---one must ensure that all devices can communicate with the controller at all times, even under arbitrary combinations of failure. Even ignoring failures, it is necessary for controller systems to
scale to meet the demands of large, geographically distributed networks,
and to react quickly
to a changing environment. To address this challenge, researchers are exploring
multi-controller systems with interacting controllers, thus bringing back distributed
control planes~\cite{mccauley2013extending,onos} and their current programming difficulties.
%However, academic
%language design and implementation efforts have not kept pace.  For instance, work on many
%experimental SDN languages~\cite{frenetic,flowlog,vericon,merlin,netkat,kinetic,pga} has not yet shown how to implement fault tolerant
%multi-controller systems that support their high-level abstractions efficiently.

Hence, in this paper, we have two central goals:
\begin{enumerate}
\item Design a new, high-level language with natural abstractions
for expressing intra-domain routing, inter-domain
routing and routing alternatives in case of failures.
\item Define algorithms for compiling these specifications into
configurations for devices running standard
distributed control plane algorithms, while ensuring correct behavior
independent of the number of faults.
\end{enumerate}

To achieve the first goal, we borrow the idea of using regular
expressions to specify network paths from 
recent high-level SDN languages such as FatTire~\cite{fattire},
Merlin~\cite{foster:merlin}, and
NetKAT~\cite{netkat}.  However, our design also contains several key
departures from existing languages.  The most important difference is semantic:  the paths specified
by programmers can extend from outside the operator's network to inside
the network, across several devices internally, and then out again.  This design
choice allows users to specify preferences about both external and internal
routes in the exact same way.
%No SDN programming language provides similar facilities.  
In addition, we augment the algebra
of regular expressions to directly support a notion of {\em preferences} and provide a semantics in terms of sets of ranked paths.  The preferences indicate fail-over behaviors:  among all specified paths that are still available,
the system guarantees that the highest-ranked one will be taken.  
%We also support network operators by defining a
%set of common abbreviations for entering, leaving, and traversing
%the user network and we provide abstractions for managing common features %of
%control plane algorithms such as route aggregation.  Operators use familiar logical operations
%to build routes in a modular and compositional manner from these
%primitives.  Our abbreviations are translated in to our regular-expression-based
%intermediate representation and compiled from there.

To achieve the second goal, we develop program analysis and compilation
algorithms that translate the regular policies to a graph-based
intermediate language and from there to BGP configurations, including
per-device import and export filters, local preferences, MED attributes, and community tags.
We target BGP for pragmatic reasons:  it is an industry
standard and many networks use it internally as well as externally. Despite the advent of SDN, many networks will continue to
use BGP for the foreseeable future due to existing infrastructure investments, the difficulty of transitioning to SDN,
and the scalability and fault-tolerance advantages of a distributed
control plane.

The BGP configurations produced by our compiler are
guaranteed to satisfy the given policy specification in the face of
{\em arbitrary} failures.  In this way, we provide network operators
with a strong guarantee that is otherwise impossible to achieve
today. 
However, some policies simply cannot be implemented correctly in BGP in 
 the presence of arbitrary failures.  We develop new
algorithms to detect such policies
and report our findings to the operators, so they may fix the policy
specification at compile time rather than experience undesirable
behavior after the configurations are deployed.
%We also analyze route aggregation to detect the possibility of failure-induced black-holes.
